---
title: "HypeRSpark vignette"
author: "Jarno E. Smit"
package: "HypeRSpark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HypeRSpark-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F}
library(hypeRSpark)
```

# Introduction

## Installing the package
Can be installed directly from GitHub, using the *devtools* package. Building the vignettes relies on the *markdown* package.
The code below illustrates how the packages can be downloaded and installed from GitHub.

```{r, eval=F}
install.packages("devtools")
install.packages("rmarkdown")
devtools::install_github("jarnos97/rPackageHypeRSpark", build_vignettes = TRUE)
```


Why was the package created.
What is the use?
Which functions are there?


# Functions

## Install dependencies

HyperSpark has the following external dependencies:

* Java SE Development Toolkit (JDK): the package was build using version 8 - update 281. 
However, it keeps compatibility with newer versions. This requires a (free) Oracle account.
Download links [here.](https://www.oracle.com/nl/java/technologies/javase/javase-jdk8-downloads.html)
* Apache Maven: the package was build using version 3.8.1. Download links [here.](https://maven.apache.org/download.cgi)
However, it keeps compatibility with newer versions.

Users should install these dependencies and add both Java and Maven to the system path. 
[This tutorial](https://javatutorial.net/set-java-home-windows-10) shows how to add Java to the system path for Windows 10,
the same method is used for Maven.


## Configuring hyperspark


## packageHyperSpark

The first time running this command can take a while, as all dependencies need  to be downloaded
from the Maven repository. Subsequent packaging is much faster. 



